---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%",
  warning = FALSE,
  message = FALSE
)
```

# rusleR

<!-- badges: start -->
<!-- badges: end -->

This package offers an `R` implementation of Universal Soil Loss Equation ([USLE](https://en.wikipedia.org/wiki/Universal_Soil_Loss_Equation)). You can find here a collection of functions to estimate main factors: R-factor, K-factor, LS-factor and C-factor. The package uses[ `terra`](https://github.com/rspatial/terra) and [`Rsagacmd`](https://github.com/stevenpawley/Rsagacmd/) in the background. [`SAGA GIS`](https://sourceforge.net/projects/saga-gis/) needs to be install on your machine as well. 

## Installation

You can install the development version of rusleR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atsyplenkov/rusleR")
```

## Example

```{r example}
library(rusleR)
## basic example code
```

This is a basic example which shows you how to calculate LS_alpine ([Schmidt et al., 2019](https://www.sciencedirect.com/science/article/pii/S2215016119300056)):

```{r ls_alpine}
library(Rsagacmd)
library(terra)

# initiate a saga object
saga <- saga_gis(raster_backend = "terra")

# load DEM
f <- system.file("extdata/dem.tif", package="rusleR")
DEM <- rast(f)

# calculate LS-alpine
ls <- ls_alpine(dem = DEM)
```

As a result of `ls_alpine()` you receive a `SpatRaster` object:
```{r summary}
ls
```

```{r pressure, echo = FALSE}
plot(ls)
```

## To add
- K-factor functions
- R-factor function to crop GLOREDA
- C-factor???
